\nwfilename{environment.nw}\nwbegindocs{0}% -*- mode: Noweb; noweb-code-mode: c-mode -*-% ===> this file was generated automatically by noweave --- better not edit it

%**********************************************************************
\chapter{Environment: liaison des addresses (noms) au dimensions (entiers)}
\label{chap:env}
%**********************************************************************

Ce fichier définit les environnements qui font le lien avec d'une
part des addresses (des noms) et d'autre part leur type et la
dimension qui leur est associée au niveau 0.

On distingue deux catégories de fonctions:
\begin{itemize}
\item Celles utiles à un utilisateur du niveau 1 de l'interface,
  pour manipuler des environnements.  Il s'agit des fonctions
  {\Tt{}environment{\_}free,\ environment{\_}make,\ environment{\_}add,\ environment{\_}remove,\ environment{\_}is{\_}eq,\ environement{\_}is{\_}leq,\ environment{\_}compare}.
\item Celles utilisées essentiellement par l'interface commune
(module abstract1) pour passer automatiquement du niveau 1 au
niveau 0, qui sont d'un usage interne
\end{itemize}


\nwenddocs{}\nwbegincode{1}\moddef{environment.h}\endmoddef\nwstartdeflinemarkup\nwenddeflinemarkup
/* ********************************************************************** */
/* environment.h: binding of addresses (strings) to dimensions */
/* ********************************************************************** */

/* GENERATED FROM environment.nw: DOT NOT MODIFY ! */

#ifndef _ENVIRONMENT_H_
#define _ENVIRONMENT_H_

#include "manager.h"
#include "expr0.h"
\nwendcode{}\nwbegindocs{2}\nwdocspar

%======================================================================
\section{Datatypes}
%======================================================================

\nwenddocs{}\nwbegincode{3}\moddef{environment.h}\plusendmoddef\nwstartdeflinemarkup\nwenddeflinemarkup

/* ====================================================================== */
/* Datatype */
/* ====================================================================== */

/* An environement is parametrized by the abstract type var_t, 
   equipped with a total ordering function, a copy function, 
   and a free function.

   The parametrization is done via a global variable pointing to an 
   environment_var_t structure.
 */

typedef void* var_t;

typedef struct environment_var_t \{
  int (*compare)(var_t v1, var_t v2);  /* Total ordering function */
  var_t (*copy)(var_t var);      /* Duplication function */
  void (*free)(var_t var);            /* Deallocation function */
  char* (*to_string)(var_t var); /* Conversion to a dynamically allocated string */
  void (*fprint)(FILE* stream, var_t var); /* Printing to a stream */
\} environment_var_t;

extern environment_var_t environment_var_default;
  /* default manager: var_t are char* */

extern environment_var_t* environment_var;

\nwendcode{}\nwbegindocs{4} 

\nwenddocs{}\nwbegincode{5}\moddef{environment.h}\plusendmoddef\nwstartdeflinemarkup\nwenddeflinemarkup
/* The environment_t type is the type used by the level 1 of the interface.
   It should be considered as an abstract type.

   We use reference counting to manage memory. Conventions are:

   - a newly allocated environement has a reference count of one;
   - environement_copy increments the counter and return its argument
   - environement_free decrements it and free the environement
     in case of zero or negative number.
*/

typedef struct environment_t \{
  var_t* var_of_dim;
  /*
    Array of size intdim+realdim, indexed by dimensions.
    - It should not contain identical strings..
    - Slice [0..intdim-1] is lexicographically sorted,
      and denotes integer variables.
    - Slice [intdim..intdim+realdim-1] is lexicographically sorted,
      and denotes real variables.
    - The memory allocated for the variables are attached to the structure
      (they are freed when the structure is no longer in use)
  */
  size_t intdim; /* Number of integer variables */
  size_t realdim;/* Number of real variables */
  size_t count; /* For reference counting */
\} environment_t;

typedef struct environment_name_of_dim_t \{
  size_t size;
  char* p[0]; /* Assumed to be of size size */
\} environment_name_of_dim_t;
\nwendcode{}\nwbegindocs{6}\nwdocspar

%======================================================================
\section{Memory management, Constructor, Destructors}
%======================================================================

\nwenddocs{}\nwbegincode{7}\moddef{environment.h}\plusendmoddef\nwstartdeflinemarkup\nwenddeflinemarkup

/* ====================================================================== */
/* Memory management, Constructor, Destructors */
/* ====================================================================== */

void environment_free2(environment_t* e);
  /* Free the environement
     (the structure itself and the memory pointed to by fields) */

static inline
void environment_free(environment_t* e);
  /* Decrement the reference counter and possibly deallocate */

static inline
environment_t* environment_copy(environment_t* e);
  /* Copy */

void environment_fdump(FILE* stream, environment_t* env);
  /* Print an environement under the form:
      environment: dim = (..,..), count = ..
      0: name0
      1: name1
      ...
  */

environment_name_of_dim_t* environment_name_of_dim_alloc(environment_t* e);
void environment_name_of_dim_free(environment_name_of_dim_t*);

environment_t* environment_make_empty();
  /* Build an empty environement */

environment_t* environment_make(var_t* name_of_intdim, size_t intdim,
                                var_t* name_of_realdim, size_t realdim);
  /* Build an environement from an array of integer and
     an array of real variables.
     - Names are duplicated in the result, so it is the responsability
       of the user to free the names he provides.
     - If the result does not satisfy the invariant, return NULL.
  */

environment_t* environment_add(const environment_t* env,
                               var_t* name_of_intdim, size_t intdim,
                               var_t* name_of_realdim, size_t realdim);
  /* Add variables to an environement.
     Same comments as for environement_make */
environment_t* environment_remove(const environment_t* env,
                                  var_t* name_of_intdim, size_t intdim,
                                  var_t* name_of_realdim, size_t realdim);
  /* Remove variables from an environement.
     Same comments as for environement_make */

dim_t environment_dim_of_var(const environment_t* env, var_t name);
  /* - If the variable is present in the environemnt,
       return its associated dimension.
     - Otherwise, return DIM_MAX
  */

static inline
var_t environment_var_of_dim(const environment_t* env, dim_t dim);
  /* - Return the variable asosciated to the dimension.
     - There is no bound check here */
\nwendcode{}\nwbegindocs{8}\nwdocspar

%======================================================================
\section{Tests}
%======================================================================

\nwenddocs{}\nwbegincode{9}\moddef{environment.h}\plusendmoddef\nwstartdeflinemarkup\nwenddeflinemarkup

/* ========================================================================= */
/* Tests */
/* ========================================================================= */

bool environment_is_eq(const environment_t* env1,
                       const environment_t* env2);
  /* Equality test */
bool environment_is_leq(const environment_t* env1,
                       const environment_t* env2);
  /* Inclusion test */
int environment_compare(const environment_t* env1,
                        const environment_t* env2);
  /* Return
    - -2 if the environement are not compatible
      (a variable has different types in the 2 environements)
    - -1 if env1 is a subset of env2
    - 0 if equality
    - +1 if env1 is a superset of env2
    - +2 otherwise (the lce exists and is a strict superset of both)
  */
\nwendcode{}\nwbegindocs{10}\nwdocspar

%======================================================================
\section{Least common environments and conversion permutations}
%======================================================================

\nwenddocs{}\nwbegincode{11}\moddef{environment.h}\plusendmoddef\nwstartdeflinemarkup\nwenddeflinemarkup

/* ========================================================================= */
/* Least common environments and conversion permutations */
/* ========================================================================= */

dimchange_t* environment_dimchange(const environment_t* env1,
                                   const environment_t* env);
  /* Compute the transformation for converting from an environment to a
     superenvironment.
     Return NULL if env is not a superenvironment.
  */

environment_t* environment_lce(const environment_t* env1,
                               const environment_t* env2,
                               dimchange_t** dimchange1,
                               dimchange_t** dimchange2);
  /*
  Least common environment to two environment.

   - Assume environment_is_eq(env1,env2)==false.

   - If environments are not compatible
     (a variable has different types in the 2 environements),
     return NULL

   - Compute also in dimachange1 and dimchange2 the conversion transofrmations.

   - If no dimensions to add to env1, this implies that env is
     actually env1. In this case, *dimchange1==NULL. Otherwise, the
     function allocates the dimchange1 with malloc.
 */
environment_t* environment_lce_array(const environment_t** tenv,
                                     size_t size,
                                     dimchange_t*** ptdimchange);
/*
  Least common environment to an array environments.

  - Assume the size of the array is at least one.

  - If all input environments are the same, *ptdimchange==NULL.
    Otherwise, compute in *ptdimchange the conversion permutations.

  - If no dimensions to add to env[i], this implies that env is actually
    env[i]. In this case, (*ptdimchange)[i]==NULL. Otherwise, the function allocates
    the array (*ptdimchange)[0] with malloc.

 */
\nwendcode{}\nwbegindocs{12}\nwdocspar

%======================================================================
\section{Renommage de variables}
%======================================================================

\nwenddocs{}\nwbegincode{13}\moddef{environment.h}\plusendmoddef\nwstartdeflinemarkup\nwenddeflinemarkup
/* ========================================================================= */
/* Variable renaming */
/* ========================================================================= */

environment_t* environment_rename(environment_t* env,
                                  var_t* tvar1, var_t* tvar2, size_t size,
                                  dim_t** permutation);
  /* Rename the variables in the environment.
     size is the common size of tvar1 and tvar2,
     and permutation is a result-parameter

     The function applies the variable substitution tvar1[i]->tvar2[i] to
     the environment, and returns the resulting environment and 
     the allocated transformation permutation in *permutation.
     
     If the parameters are not valid, returns NULL with *permutation==NULL
  */
\nwendcode{}\nwbegindocs{14}\nwdocspar

%======================================================================
\section{Définitions en-ligne}
%======================================================================

\nwenddocs{}\nwbegincode{15}\moddef{environment.h}\plusendmoddef\nwstartdeflinemarkup\nwenddeflinemarkup

/* ====================================================================== */
/* Inline definitions */
/* ====================================================================== */

static inline
var_t environment_var_of_dim(const environment_t* env, dim_t dim)\{
  return dim < env->intdim+env->realdim ? env->var_of_dim[dim] : NULL;
\}
static inline
void environment_free(environment_t* env)\{
  if (env->count<=1)
    environment_free2(env);
  else
    env->count--;
\}
static inline
environment_t* environment_copy(environment_t* env)\{
  env->count++;
  return env;
\}
\nwendcode{}\nwbegindocs{16}\nwdocspar

\nwenddocs{}\nwbegincode{17}\moddef{environment.h}\plusendmoddef\nwstartdeflinemarkup\nwenddeflinemarkup
#endif
\nwendcode{}\nwbegindocs{18}\nwdocspar
\nwenddocs{}
